---
title: "Fundamentals 3: EDA & Data Visualization"
output: html_notebook
---

**Objective**: Learn about and apply different plotting methods using the `iris` dataset. 

# Install and load necessary packages

```{r Intall packages, include = F}
# ggplot2
  install.packages('ggplot2')
  library(ggplot2)

# ggally
  install.packages('ggally')
  library(ggally)
  
# datasets (iris)
  library(datasets)
```

# Load the `iris` dataset using the `data` function

```{r Load iris dataset}

```

# Pair plot

**Objective**: Generate pair plots for all variables in the `iris` dataset.

```{r Visualize pair plots}

```

# Violin plot

**Objective**: Generate violin plots to compare `Petal.Length` between different species. 

```{r Visualize violin plots}

```

# Parallel coordinate plot

**Objective**: Generate parallel coordinate plots to identify any trends across variables for each species. 

```{r Visualize parallel coordinate plots}

```

# Heatmap

**Objective**: Generate a heatmap for `Petal.Length` to determine whether this feature can differentiate species. 

```{r Visualize heatmap}

```
